---
output:
  word_document: default
  html_document: default
---
Libraries
```{r = FALSE}
library("tidyverse")
library("tidymodels")
```

```{r}
trucks <- read_csv("trucks.csv")
```

Task 1
```{r}
ggplot(trucks, aes(Distance, Speeding)) + geom_point()
```

There does look to be some obvious natural clustering of drivers in relation to the speeding variable. 

Task 2
```{r}
kmeans_recipe = recipe(~ Speeding + Distance, trucks)

trucks_dummy = kmeans_recipe %>%
  step_scale(all_numeric()) %>%
  step_center(all_numeric())

trucks_dummy = prep(trucks_dummy, trucks)

trucks_cleaned = bake(trucks_dummy, trucks)
```

Task 3
```{r}
set.seed(64)
clusts = 
  tibble(k = 2) %>%
  mutate(
    kclust = map(k, ~kmeans(trucks_cleaned, .x)),
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, trucks_cleaned)
  )

clusts
```

```{r}
clusters = 
  clusts %>%
  unnest(cols = c(tidied))

assignments = 
  clusts %>% 
  unnest(cols = c(augmented))

clusterings = 
  clusts %>%
  unnest(cols = c(glanced))

```

```{r}
p1 = 
  ggplot(assignments, aes(x = Distance, y = Speeding)) +
  geom_point(aes(color = .cluster), alpha = 0.8) + 
  facet_wrap(~ k)
p1
```
There are two natural clusters evenly distributed between 1 and 2.

Task 4
```{r}
set.seed(412)
clusts = 
  tibble(k = 1:8) %>%
  mutate(
    kclust = map(k, ~kmeans(trucks_cleaned, .x)),
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, trucks_cleaned)
  )

clusts

```

```{r}
clusters = 
  clusts %>%
  unnest(cols = c(tidied))

assignments = 
  clusts %>% 
  unnest(cols = c(augmented))

clusterings = 
  clusts %>%
  unnest(cols = c(glanced))
```

```{r}
p2 = 
  ggplot(assignments, aes(x = Distance, y = Speeding)) +
  geom_point(aes(color = .cluster), alpha = 0.8) + 
  facet_wrap(~ k)
p2
```
A k value of 4 or 5 looks to be most appropriate.

Task 5
```{r}
ggplot(clusterings, aes(k, tot.withinss)) + 
  geom_line() +
  geom_point()
```

4 appears to be the best value of k as that is where the values start to level out (it is the bend in the elbow).

Task 6
```{r}
set.seed(64)
clusts = 
  tibble(k = 4) %>%
  mutate(
    kclust = map(k, ~kmeans(trucks_cleaned, .x)),
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, trucks_cleaned)
  )

clusts

```

```{r}
clusters = 
  clusts %>%
  unnest(cols = c(tidied))

assignments = 
  clusts %>% 
  unnest(cols = c(augmented))

clusterings = 
  clusts %>%
  unnest(cols = c(glanced))
```

```{r}
p3 = 
  ggplot(assignments, aes(x = Distance, y = Speeding)) +
  geom_point(aes(color = .cluster), alpha = 0.8) + 
  facet_wrap(~ k)
p3
```

These resulting clusters are much better than the previous ones. There are no outlier variables and all clusters are well defined. 